### 2019-12-10 Paper/__latexindent_temp.tex

\begin{itemize}
    \item Usage trait databases: 
    \begin{itemize}
    \item

    \end{itemize}
\end{itemize}