### 2020-08-31 Paper/Flowchart/Flowchart_methods.tex
31.08: Intro+Methods+Results update

\documentclass[../Draft_harmonization_paper.tex]{subfiles}
\usetikzlibrary{trees,
                arrows,
                shapes.geometric,
                positioning,
                calc,
                backgrounds,
                fit,
                matrix,
                shapes.geometric}

\begin{document}

\tikzset{
	font = {\sffamily},
	intstepsnode/.style = {rectangle, draw = black, fill = gray!50, minimum width = 4cm, minimum height = 1cm, text centered, rounded corners, very thick},
  aggrtypesnode/.style = {rectangle, draw = black, fill = blue!25, minimum width = 4cm, minimum height = 1cm, text centered, rounded corners, very thick},
  aggrstepsnode/.style = {rectangle, draw = black, fill = purple!40, minimum width = 4cm, minimum height = 1cm, text centered, rounded corners, very thick},
  mainnode/.style = {rectangle, draw = black, fill = orange!50, minimum width = 4cm, minimum height = 1.5cm, text centered, rounded corners, very thick},
  frame/.style = {draw, rectangle, rounded corners, dashed, black, line width = 1pt, inner sep = 2.5mm},
  roundnode/.style={circle, draw=green!60, fill=green!5, very thick, minimum size=7mm},
  arrow1/.style = {->, > = stealth, draw = black, color = gray!80, line width = 2.5pt}
}

\begin{tikzpicture}
      %% Nodes
      \node(duplTaxa)[intstepsnode, align = center] at (-5cm,0cm){Allocating duplicate taxa \\ NOA \& AUS DB};
      \node(normalAUS)[intstepsnode, align = center, xshift = 0cm, yshift = 0cm]{AUS traits standardized to [0-1] \\ because multiple codings used \\ throughout this database};
      \node(omission)[intstepsnode, align = center, xshift = 5cm, yshift = 0cm]{Omitted taxa lower \\ family-level};

      \node (frameProcessing) [frame, fit = (duplTaxa) (normalAUS) (omission) ] {};
      \node (frameText) [above, black, inner sep = 2mm] at (frameProcessing.north){Data preprocessing};

      \node(harmonization)[mainnode, align = center, xshift = -5cm, yshift = -2.5cm]{Harmonization};

      \node(conversionPerc)[intstepsnode, align = center, xshift = 0cm, yshift = -2.5cm]{Traits converted to percentages \\ and normalized to [0-1]};

      \node(traitAgg)[mainnode, align = center, xshift = 5cm, yshift = -2.5cm]{Trait aggregation};

      \node(frameHarmAggr) [frame, fit = (harmonization) (traitAgg)] {};

      \node(tab_aggregation)[xshift = -3cm, yshift = -6.6cm]{
        \begin{tabular}{ccc|ccc}
        Family & Genus & Species & \specialcell{Trait \\ A\_h\textsubscript{1}} & \specialcell{Trait \\ A\_h\textsubscript{2}} & \specialcell{Trait \\ A\_h\textsubscript{3}} \\
        \hline
         F\textsubscript{1} & G\textsubscript{1} & S\textsubscript{1} & 0.25 & 0.75 & 0\\
         F\textsubscript{1} & G\textsubscript{1} & S\textsubscript{2} & 0.5 & 0.5 & 0\\
         F\textsubscript{1} & G\textsubscript{1} & S\textsubscript{3} & 0.5 & 0.5 & 0 \\
         F\textsubscript{1} & G\textsubscript{2} & S\textsubscript{4} & 0 & 1.0 & 0 \\
         F\textsubscript{1} & G\textsubscript{2} & \textit{NA} & 0.25 & 0.75 & 0 \\
         F\textsubscript{1} & \textit{NA} & \textit{NA} & 0.25 & 0.75 & 0 \\
      \end{tabular}};

      \node(directAgg)[aggrtypesnode, align = center, xshift = 4cm, yshift = -5cm]{Direct aggregation};

      \node(tab_direct_agg)[xshift = 4cm, yshift = -7.5cm]{
        \begin{tabular}{c|ccc}
        Family & Trait & Method & Value \\
        \hline
        F\textsubscript{1} & A\_h\textsubscript{1} & mean & 0.29 \\
        F\textsubscript{1} & A\_h\textsubscript{2} & mean & 0.71 \\
        F\textsubscript{1} & A\_h\textsubscript{3} & mean & 0 \\
        \hline
        F\textsubscript{1} & A\_h\textsubscript{1} & median & 0.25 \\
        F\textsubscript{1} & A\_h\textsubscript{2} & median & 0.75 \\
        F\textsubscript{1} & A\_h\textsubscript{3} & median & 0 \\
      \end{tabular}};

      \node(stepwiseAgg)[aggrtypesnode, align = center, xshift = -3.5cm, yshift = -10.3cm]{Stepwise aggregation};

      \node(tab_stepwise_agg)[xshift = -3.5cm, yshift = -12.5cm]{
        \begin{tabular}{cc|ccc}
        Family &Genus & Trait & Method & Value \\
        \hline
        F\textsubscript{1}& G\textsubscript{1} & A\_h\textsubscript{1} & mean & 0.4167\\
        F\textsubscript{1}& G\textsubscript{2} & A\_h\textsubscript{1} & mean & 0.125\\
        F\textsubscript{1}& NA & A\_h\textsubscript{1} & - & 0.25\\
      \end{tabular}};

      \node(stepwiseAgg2)[aggrstepsnode, align = center, xshift = -3.5cm, yshift = -14.7cm]{ $A\_h_1 = mean(0.4167, 0.125, 0.25)$ \\ $ = 0.26$};

      \node(weightedAgg)[aggrtypesnode, xshift = 4cm, yshift = -10.3cm]{Weighted aggregation};

      \node(weights)[aggrstepsnode, align = center, xshift = 4cm, yshift = -12.7cm]{Weights based on \\ species per genera: \\
      $3$ for $G_1$, $1$ for $G_2$, \\
      weight for $F_1$ set to $1$};

      \node(weights2)[aggrstepsnode, align = center, xshift = 4cm, yshift = -15.1cm]{$A\_h_1 = \frac{0.25*3+0.5*3+...+ 0.25*1}{3+3...+1}$ \\ \\ $ = 0.35$};

      \node(frameExample) [frame, fit = (tab_aggregation) (directAgg)(tab_direct_agg) (weights2)] {};

      %%lines
      \draw[->] (duplTaxa.east) -- (normalAUS.west);
      \draw[->] (normalAUS.east) -- (omission.west);
      \draw[->] (harmonization.east) -- (conversionPerc.west);
      \draw[->] (conversionPerc.east) -- (traitAgg.west);
      \draw[arrow1] (traitAgg.south) -- (frameExample.north);
      \draw[->] (directAgg.south) -- (tab_direct_agg.north);
      \draw[->] (stepwiseAgg.south) -- (tab_stepwise_agg.north);
      \draw[->] (tab_stepwise_agg.south) -- (stepwiseAgg2.north);
      \draw[->] (weightedAgg.south) -- (weights.north);
      \draw[->] (weights.south) -- (weights2.north);
\end{tikzpicture}

\end{document}