### 2020-10-16 Paper/Results/Results_trait_harmonization.tex
16.10: Paper Update, first implm. of Co-Authors

\documentclass[../Draft_harmonization_paper.tex]{subfiles}
% \documentclass{article}
% \usepackage[utf8]{inputenc}
% \usepackage{hyperref}
% \usepackage{float}
% \usepackage[table,xcdraw]{xcolor}
% \usepackage{color, colortbl}
% \usepackage{longtable}
% %\usepackage[sort&compress,square,comma,authoryear]{natbib}
% \usepackage{booktabs}
% \usepackage{graphicx}
% \graphicspath{{/home/kunz/Dokumente/Projects/Trait_DB/Invertebrate_traits/Paper/Figures/}}
% \usepackage{longtable}
% \usepackage{rotating}
% \usepackage{geometry}
% \usepackage{array}

% \definecolor{Gray}{gray}{0.9}

% \newcommand{\specialcell}[2][c]{%
%   \begin{tabular}[#1]{@{}c@{}}#2\end{tabular}}

\begin{document}
%! Trait (species) scores along segments of river

\subsection*{Re-analysis of Szöcs et al. using harmonized and aggregated grouping features}

Overall, using the harmonized grouping features led to only slightly different results in comparison to the original analysis (Figure \ref{fig:rda_species_scores} and SI). According to the RDA of the trait composition, sites with high salinity were characterized by families that possess the traits multivoltinism, ovoviviparity, gill-respiration, and feeding mode shredder. Only families with the trait life cycle duration $> 1$ year failed to characterize sites with high salinization. Also, life cycle duration $<= 1$ year did not characterise sites not impacted by salinity. Like in the original analysis, transition and upstream sites from the point source were characterised by univoltine taxa and taxa that lay their eggs in aquatic environments. We also constructed the linear models of the original analyses, using the traits on the extremes of the conductivity axis and found results similar to the original analyses (Table S\ref{stab:linear_models_new} and S\ref{stab:linear_models_edi}).

For every aggregation method compared, using at family-level aggregated traits did only slightly change the species scores compared to not aggregated traits (Figure \ref{fig:rda_species_scores}). 
% Hence, the interpretation of the trait composition is the same as when only using harmonized grouping features. 
%? Mention that threshold according to mahalanobis distance is almost reached for life cycle duration

% Species scores
\begin{figure}[H]
    \label{fig:rda_species_scores}
    \centering
    \includegraphics[width=16.5cm, height=10cm]{Species_scores_rda.png}
    \caption{Species scores obtained by RDA from the original analysis \cite{szocs_effects_2014}, using harmonized grouping features, and using harmonized grouping features with assigned trait affinities aggregated to family-level.}
    \label{fig:rda_species_scores}
\end{figure}

% PUP comment: 
% What are exactly „RDA species scores“? The coordinates of species along the first factorial axis in RDA?
% If „yes“, I’m not sure this result presentation is suitable. In fact the box-plots summarizing the distributions of the set of species in the RDA analysis, after applying different strategies of trait description, can be very similar (it’s not surprising, because in each case the mean coordinate of species is 0), but the individual locations of species along each axis can be very different from an RDA to another axis (i.e. from a grouping feature definition strategy to another). It would be probably far more interesting to represent the distribution of deviations between the RDA scores provided by each strategy and the original RDA scores (from Szöcs et al., 2018)..

\end{document}