### 2021-02-09 Paper/Draft_harmonization_paper.tex
09.02: Brook Edits part 3

\documentclass{article}
\usepackage[utf8]{inputenc}
\usepackage{hyperref}
\usepackage{float}
\usepackage[table,xcdraw]{xcolor}
\usepackage{color, colortbl}
\usepackage{longtable}
%\usepackage[sort&compress,square,comma,authoryear]{natbib}
\usepackage{booktabs}
\usepackage{graphicx}
\usepackage{multirow}
\usepackage{tikz}
\usepackage{rotating}
\usepackage{caption}
% \graphicspath{{/home/kunz/Dokumente/Projects/Trait_DB/Invertebrate_traits/Paper/Figures/}}
\graphicspath{{Figures/}}
\usepackage{rotating}
\usepackage{geometry}
\usepackage{array}
\usepackage{lscape}
\usepackage{longtable}
\usepackage{hhline}
\usepackage{lmodern}
\usepackage[
backend = biber,
natbib,
style = apa,
citestyle = authoryear,
maxcitenames = 2, 
maxbibnames = 99, 
uniquename=false,
uniquelist=false
%sorting = none % sort by name year title
]{biblatex}
\addbibresource{Ref_invertebrate_DB.bib}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% From freshwater biology:
% List all sources in the reference list alphabetically by name. In text citations should follow the author-date method. This means that the author's last name and the year of publication for the source should appear in the text, for example, (Jones, 1998), and a complete reference should appear in the reference list at the end of the paper.

% References are styled according to the sixth edition of the Publication Manual of the American Psychological Association. A sample of the most common entries in reference lists appears below. Please note that for journal articles, issue numbers are not included unless each issue in the volume begins with page one.

% Article: 
% One author: Fawcett, T. (2006). An introduction to ROC analysis. Pattern Recognition Letters, 27(8), 861–874. DOI: 10.1016/j.patrec.2005.10.010.

% 2 to 7 authors: Daley, C. E., & Nagle, R. J. (1996). Relevance of WISC-III Indicators for assessment of learning disabilities. Journal of Psychoeducational Assessment, 14 (4), 320–333.

% More than 7 authors: Rutter, M., Caspi, A., Fergusson, D., Horwood, L. J., Goodman, R., Maughan, B., … Carroll, J. (2004). Sex differences in developmental reading disability: New findings from 4 epidemiological studies. Journal of the American Medical Association, 291(16), 2007–2012. DOI: 10.1001/jama.291.16.2007

% In press or forthcoming: van Bergen, E., de Jong, P. F., Maassen, B., Krikhaar, E., Plakas, A., & van der Leij, A. (in press). IQ of four-year-olds who go on to develop dyslexia. Journal of Learning Disabilities. DOI: 10.1177/0022219413479673
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\usepackage{subfiles} % Best loaded last in the preamble

% functions and definitions
\definecolor{Gray}{gray}{0.9}

% horizontal space between two columns
\setlength{\tabcolsep}{2mm}

\newcommand{\specialcell}[2][c]{%
  \begin{tabular}[#1]{@{}c@{}}#2\end{tabular}}

\renewcommand*{\thefootnote}{\alph{footnote}}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\title{Tackling discrepancies in freshwater invertebrate trait databases: Harmonising across continents and aggregating taxonomic resolution}
\author{}%Stefan Kunz 
\date{}

\begin{document}
\maketitle

\section*{Abstract}

% FROM FRESHWATER BIOLOGY Website:
% All papers should include a summary, in short numbered paragraphs, limited to about 3% of the length of the text, and in any case to not more than 500 words. This should provide a concise statement of the scope of the work and its principal findings and be fully intelligible without reference to the main text. It is helpful to structure your abstract as follows: point 1 = background to and reason for the study and the aims; point 2 = brief description of methods used; point 3 = main project results (address the aims); point 4 = conclusions of your study; point 5 = the contribution that your results make to the field / why your study is useful to an international audience. Referring to abstracts in recently published issues may help with structuring an abstract.

% Alternative start: Invertebrate traits are increasingly used in ecological studies because they provide a mechanistic understanding of species-environment relationships and are suitable for large scale analysis due to their consistent response to environmental gradients.

1. Use of invertebrate traits rather than species composition may facilitate large-scale comparisons of community structure and responses to disturbance in freshwater ecology because the same traits can potentially occur everywhere.  In recent years, comprehensive invertebrate trait databases have been established at different scales (e.g. regions, continents). The wide availability of invertebrate trait data supports trait-based studies, especially at large scales. However, a number of data-related issues complicate the use of invertebrate traits for ecological studies. For example, standardised definitions for for freshwater invertebrate traits are lacking, which impedes comparisons across regions. Moreover, it is uncertain how harmonising varying trait definitions among databases might influence identification of trait-environment relationships. In addition, is that taxonomic resolution between observational taxonomic data and trait data can differ, making aggregation of traits necessary. At present, it is unknown how different trait aggregation approaches compare with expert-assigned trait affinities. In this study we aimed to identify discrepancies in trait definitions across freshwater invertebrate databases and to compare trait aggregation methods to identify how taxonomic hierarchy structure and trait variability influence results from different aggregation approaches.

2. We describe discrepancies in the definitions of traits used to create freshwater invertebrate trait databases in Europe, North America, New Zealand, and Australia. Based on our comparisons of these trait databases, we established four novel trait datasets by harmonising trait definitions of commonly used traits. Next, we used two of these datasets to compare aggregated traits obtained by different aggregation methods with traits assigned by experts, both at the family level. The trait aggregation methods we compared used either the mean or the median and different weightings. %Thereby, we simulated how taxonomic resolution and trait variability can change the result of different aggregation approaches.
We further explored the effects of harmonisation and trait aggregation by re-analysing data from a case study.

3. We found that among trait databases, trait definitions often differed because varying numbers of traits were used to describe groups of related traits (e.g., respiration traits) and the focus of descriptions for groups of related traits also varied (e.g., for feeding mode some databases focused on the food source, whereas others focused on mouthpart morphology). The coding to describe traits (binary, fuzzy) also varied among databases. Our comparison of different aggregation methods showed that family-level aggregated and expert-assigned traits were similar, especially from approaches using the median, and that trait coding method, uneven taxonomic hierarchical structure, and high trait variability can influence aggregated trait affinities. We further showed that harmonised and aggregated data performed similarly to observational data in identifying trait–environment relationships.

4. By outlining discrepancies in trait definitions we hope to motivate the development of standardised terminology for invertebrate traits. Our results also illustrate the usefulness of harmonised datasets for ecological study and provide guidance for the circumstances under which the choice of trait aggregation method is important.

\newpage

\section*{Introduction}

Explaining and predicting how communities are shaped by environmental factors is a primary main goal of ecology. Species traits are measurable properties of an organism (\cite{mcgill_rebuilding_2006}), and comparing communities based on the traits possessed by species may facilitate testing of a range of ecological hypotheses (\cite{heino_jani_macroecological_2013}). Traits are assumed to be adaptations (e.g., physiological, behavioural) of organisms to their environment and to indicate direct or indirect linkages between the biological response of an organism or a population and its environment (\cite{southwood_habitat_1977, verberk_delivering_2013}).
In addition to providing local-scale mechanistic-based expressions of species-environment relationships, trait-based approaches might be suitable for large-scale analyses because trait responses are less constrained by biogeographic boundaries and taxon distributional areas than taxonomic responses (\cite{baird_toward_2011, bonada_taxonomic_2007}).

Invertebrate traits have been increasingly used in freshwater ecology, e.g. by relating macroinvertebrate trait composition to environmental factors (\cite{bhowmik_large_2015, poff_developing_2010, szocs_effects_2014}). In the last decades, freshwater ecologists have compiled comprehensive invertebrate trait databases, which typically include data for a single region or continent (\cite{kefford_integrated_2020, Philips_and_Smith_NZ_DB_2018, schmidt-kloiber_www.freshwaterecology.info_2015, tomanova_trophic_2006, ussegliopolatera_biological_2000, vieira_database_nodate}). The availability of invertebrate trait data from different continents enables comparisons of regional trait variation and allows for testing the consistency of trait structure–environmental factor relationships across both small and large spatial extents. To date, such analyses have been carried out mostly within continents, using information from one or two trait databases. For example, \citet{bonada_taxonomic_2007} compared trait composition for Mediterranean and temperate regions in Europe based on traits from \citet{ussegliopolatera_biological_2000} (typically referred to as the Tachet database), \citet{poff_developing_2010} characterized trait composition across sites in the Western USA based on traits from \citet{poff_functional_2006}, and \citet{botwe_effects_2018} used trait definitions from \citet{poff_functional_2006} and \citet{schafer_trait_2011} to test for effects of salinity on invertebrate traits across different sites in South Australia. Analyses that synthesize invertebrate trait information from more than two different continents are rare, but see \citet{brown_functional_2018} and \citet{statzner_reproductive_1997}.

The heterogeneity of information in freshwater invertebrate trait databases, besides the diversity of taxa across regions, is likely a major reason for the lack of studies across continents. There are several challenges presented by using invertebrate trait data from multiple established databases in ecological studies. One such challenge is inconsistency in terminology. In this study, we follow the terminology proposed by \citet{schmera_proposed_2015}, where a grouping feature is defined as a general property (e.g. feeding mode) that comprises a "group of related traits (e.g., predator, shredder, etc.) that vary among species or among individuals within a species". Thus, we use the term grouping feature in place of the term trait, and the term trait instead of trait state, modality or trait category. 
% Ben suggest here: For  example  e.g., predator, shredder, etc. Are traits of the grouping feature ..... and removing predator, shredder, etc. from the sentence above.

Another challenge to using data from multiple continents is that invertebrate trait databases include different grouping features and related traits. To harmonize grouping features from different regions, or to structure them in such a way as to facilitate cross-database comparison, commonly accepted and unambiguous trait definitions are required (\cite{schneider_towards_2019}). Ideally, the same traits and grouping features would be reported across databases or standardised terminology would exist to easily harmonise them; however, there is a lack of standardised terminology of trait definitions and poor metadata quality in many trait databases, making grouping-feature harmonisation difficult (\cite{baird_toward_2011, kissling_towards_2018}). To our knowledge, only \citet{brown_functional_2018} harmonised grouping features from more than two geographically distant invertebrate trait databases, for a limited set of grouping features (8) and taxa (112), in a study on the influence of decreasing glacier cover on functional diversity and invertebrate assemblages.

A third challenge to comparing across invertebrate trait databases is inconsistent coding of trait data (\cite{culp_incorporating_2011}), or assigning affinities (aka affinity scores). Traits of individual freshwater invertebrates can be difficult to quantify because, unlike plant traits, they are often difficult to measure directly. For example, to describe feeding habits requires that we understand mouthpart morphology, feeding behaviour, and specific food sources (\cite{moog_comprehensive_nodate}). Some traits can be described as binary or categorical, which ignores uncertainty in how traits are expressed in any particular organism (e.g., adult terrestrial stage, presence of gills). Other traits are better suited to being described as continuous (e.g., tolerance of pollution, body size). One approach for dealing with uncertainty is the use of fuzzy coding, where traits are assigned probabilistic values. Fuzzy codes, which are usually converted to proportions, are used to account for plasticity in traits, variability in traits within taxonomic groups above the species level, and incomplete knowledge. A challenge is presented by the use of different types of coding for trait affinities in different databases. For example, in the study from \citet{brown_functional_2018} the authors needed to reclassify trait affinities because the individual databases employed different coding approaches (i.e., European and New Zealand databases used fuzzy coding, North American database used binary coding).

Finally, discrepancies in the taxonomic resolutions (e.g., species, genus, or family) when comparing among trait databases, or when linking observational taxonomic data to trait databases, presents another challenge. When observations or data in one trait database are at a more precise taxonomic level than other data (e.g. observations at species-level and trait data at genus-level) trait information of the less precise taxonomic level is often assigned (e.g. \cite{szocs_effects_2014, vos_taxonomic_2017}). Conversely, if trait information is only available at more precise taxonomic levels than observed taxa, traits are aggregated to a less precise taxonomic level (e.g. \cite{aspin_extreme_2019, piliere_a._f._h._importance_2016, poff_functional_2006, szocs_effects_2014}). To date, studies have used different methods of trait aggregation, e.g. the mean (\cite{magliozzi_functional_2019}), median (\cite{szocs_effects_2014}) or mode (\cite{piliere_a._f._h._importance_2016}), but studies on how and to what extent different trait aggregation methods influence trait-based analyses are missing. 
%However, related knowledge would inform future studies regarding the choice of the aggregation method.

These challenges when working with or synthesizing trait data motivated us to assess the characteristics and comparability of freshwater invertebrate trait data contained in disparate databases. We aimed to explore methods of improving the comparability of trait data across databases, despite differences in trait definitions and the taxonomic level of data in three ways: (1) by investigating and describing discrepancies in trait definitions between trait databases from Europe, North America, Australia, and New Zealand; (2) by harmonising trait definitions across databases to provide consistent trait grouping features for comparison across geographic locations; and (3) by comparing the influence of different trait aggregation methods and grouping-feature harmonisation on the identification of trait–environment relationships.

\newpage
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section*{Methods}

To address our study objectives, we took a multi-prong approach to describing, combining, and assessing multiple sources of freshwater invertebrate data that differed in trait grouping-feature definitions and taxonomic resolution. 
%Version from Brooke: As a basis for evaluating the comparability of freshwater invertebrate databases, we developed four novel invertebrate trait datasets by harmonising seven grouping features from existing trait databases. We then evaluated the influence of different...
Identified trait definitions discrepancies served as a starting point to develop four novel invertebrate trait datasets by harmonising seven grouping features from existing trait databases. The harmonised trait datasets were used to evaluate the influence of different trait aggregation approaches on the identification of trait–environment relationships, our third study objective, in two ways by (1) comparing family-level aggregated trait affinities with family-level trait affinities assigned by experts for two of the established trait datasets where such data were available and (2) comparing the influence of taxonomic hierarchy and trait variability on the outcomes of the different trait aggregation approaches for simulated data. Finally, to further investigate how harmonising and aggregating trait data can modify the outcomes of trait–environment relationship characterisation, we used data from one of the newly established datasets with harmonised and aggregated traits to repeat an analysis of salinisation effects on biological traits (\cite{szocs_effects_2014}) and compared our results with those of the original study.

\subsection*{Selection of traits and harmonisation of trait databases}

To explore the characteristics and comparability of freshwater invertebrate trait data with differing grouping-feature definitions and taxonomic resolution, we obtained invertebrate trait data from six trait databases for four regions: Europe, North America, Australia, and New Zealand. For Europe we obtained trait information from the freshwaterecology.info database (\cite{schmidt-kloiber_www.freshwaterecology.info_2015}), which contains taxa at the species level, and the Tachet database (\cite{ussegliopolatera_biological_2000}), which contains taxa at species, genus, and family levels. For North America we primarily obtained trait information from \citet{twardochleb_freshwater_nodate} (hereafter referred to as the CONUS database) and where taxa were missing from that database, we obtained data from \citet{vieira_database_nodate} (hereafter referred to as the Vieira database). Both databases contain taxa at species, genus and family levels. Data on body form for European and North American taxa were based on expert knowledge (\cite{polatera_personal_information_2020}). For Australia and New Zealand, we used trait databases from \citet{kefford_integrated_2020} (hereafter referred to as the Australian database) and \citet{Philips_and_Smith_NZ_DB_2018} (hereafter referred to as the New Zealand database) respectively. These databases contain taxa at species, genus, family, and less precise taxonomic levels.

We established four datasets, one for each of the four geographic regions, by harmonising the traits of seven grouping features from the trait databases. We selected traits that were available in all databases, are commonly applied in trait-based ecological studies, and describe different parts of the biology of a species: life history (Voltinism), morphology (Respiration, Body form, Size), ecology (Locomotion, Feeding mode) and reproduction (Oviposition). We omitted ecological traits that describe habitat preferences (e.g. temperature preference) because these traits are missing in the New Zealand trait database. Grouping features were  classified differently across the databases, so we harmonised them into 26 traits (Table \ref{tab:traits_harmonisation}). To do so we combined  similar traits into one trait (e.g. crawlers and sprawlers into crawlers) and assigned the highest trait affinity score among the combined traits for each taxon.

Next, we selected taxa for inclusion in our datasets and normalised the trait affinity scores for each taxon. We first omitted all taxa with a lower taxonomic precision than family level and consolidated duplicate taxa that appeared within a database, by either applying the median for fuzzy coded traits, or the maximum for binary traits. We then established our harmonised datasets by using either fuzzy coding or, when unavailable, binary coding. Fuzzy codes are reported with different ranges in the trait databases (e.g., freshwaterecology.info: 0–10, Tachet database: 0–3 or 0–5). We normalised these to a range between 0 and 1 and converted trait affinities to proportions. In cases where binary coding was needed, we converted categorical and continuous traits into binary traits with values of 1 and 0, which, thus, corresponded with the range of fuzzy coded traits. For example, in the  freshwaterecology.info database, the classification of the grouping feature voltinism accounts for different faunistic regions and we substituted a value of 1 for entries such as “arctic” or “boreal”. We assumed that values of 1 and 0 for binary traits corresponded to the highest and no affinity for a particular trait.

\begin{table}[H]
\centering
\caption{Traits of harmonised grouping features from six invertebrate trait databases and four geographic regions. The last column indicates traits that were combined for harmonisation (no combining needed if empty).}
\label{tab:traits_harmonisation}
\begin{tabular}{lll}
\toprule[.1em]
Grouping feature & Trait & Combined traits\\
\toprule[.1em]
Voltinism    & \begin{tabular}[c]{@{}l@{}}Semivoltine\\ Univoltine\\ Bi/multivoltine\end{tabular}                                & \begin{tabular}[c]{@{}l@{}}\textless 1 generation per year\\ 1 generation per year\\ \textgreater 1 generation per year\end{tabular}                                                                                                                                            \\
\midrule
Body Form    & \begin{tabular}[c]{@{}l@{}}Cylindrical \\ Flattenend\\ Spherical\\ Streamlined\end{tabular}                       & \begin{tabular}[c]{@{}l@{}}Cylindrical, tubular\\ Flattenend, dorsoventrally flattened$^{\dagger}$ \\ Spherical, round (humped)\\ Streamlined, fusiform\end{tabular}                                                                                                                                                                          \\
\midrule
Size         & \begin{tabular}[c]{@{}l@{}}Small \\ Medium \\ Large\end{tabular}                                                  & \begin{tabular}[c]{@{}l@{}}\textless 9 mm, \textless 10 mm$^{\ddagger}$ \\ 9 - 16 mm, 10 - 20 mm\\ \textgreater 16 mm, \textgreater 20 mm\end{tabular}                                                                                                                \\
\midrule
Respiration  & \begin{tabular}[c]{@{}l@{}}Gills\\ Plastron/Spiracle\\ \\ \\ \\ Tegument\end{tabular}                             & \begin{tabular}[c]{@{}l@{}} Tracheal gills, gills\\ Temporary air store, spiracular gills, \\ atmospheric breathers, plant breathers, \\ functional spiracles, air (plants), aerial, \\ plastron/spiracle\\ Cutaneous, tegument \end{tabular}                                                                         \\
\midrule
Locomotion   & \begin{tabular}[c]{@{}l@{}}Burrower\\ Crawler\\ Sessile\\ Swimmer \end{tabular}                                     & \begin{tabular}[c]{@{}l@{}}Interstitial, boring, burrowing\\ Sprawler, walking, climber, clinger, crawler\\ Attached, sessile\\ Skating, diving, planctonic, swimming\end{tabular}                                                                                                                 \\
\midrule
Feeding mode & \begin{tabular}[c]{@{}l@{}}Filterer\\ \\ Gatherer\\ \\ Herbivore\\ \\ Parasite\\ Predator\\ Shredder \\ \\ \end{tabular} & \begin{tabular}[c]{@{}l@{}}Active/passive filterer, absorber, \\ filter-feeder, collector-filterer, filterer\\ Deposit-feeder, collector-gatherer, \\ detrivore, gatherer\\ Grazer, scraper, piercer herbivore, \\ herbivore, algal piercer, piercer (plants)$^{\mathsection}$\\ \\ Piercer (animals)$^{\mathsection}$, predator \\ Miner, xylophagus, shredder, \\ shredder detrivore\end{tabular} \\
\hline
Oviposition  & \begin{tabular}[c]{@{}l@{}}Aquatic eggs\\ \\ Ovoviviparity\\ Terrestrial eggs\end{tabular}                            & \begin{tabular}[c]{@{}l@{}}Eggs attached to substrate/plants/stones,\\ free/fixed eggs/clutches\\ \\ Terrestrial clutches, terrestrial \end{tabular}     \\
\bottomrule[.1em]
\end{tabular}
\end{table}
\begin{minipage}{\linewidth}{\fontsize{8}{10}\selectfont
    $\dagger$ The trait "bluff (blocky)" occurred in the Vieira database and was newly classified by expert knowledge into cylindrical and flattened (\cite{polatera_personal_information_2020}). \\  
    $\ddagger$ Reflects the different size classifications by the Vieira and CONUS databases from the other trait databases. \\
    $\mathsection$ The trait piercer was defined in the Tachet database for piercing plants and animals, in contrast to the other databases (\cite{usseglio-polatera_biomonitoring_2000}). Taxa exhibiting this trait have been assigned to predators or herbivores based on expert knowledge (\cite{polatera_personal_information_piercer_2020}).
}
\end{minipage}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\subsection*{Comparing trait aggregation methods}

For our study aim of comparing the effects of different aggregation methods on trait–environment relationship inference, we aggregated the traits of two of the harmonised datasets to the family level.
We used three different aggregation approaches: (1) direct aggregation of taxa to family level with either the median or mean, giving equal weight to every taxon. We denote these aggregation methods as \textit{direct\_agg\textsubscript{median}} and \textit{direct\_agg\textsubscript{mean}}, respectively; (2) stepwise aggregation, first to the genus level and subsequently to the family level with either the median or mean. This approach gives equal weights to each genus. Hereafter, we denote this aggregation type as \textit{stepwise\_agg\textsubscript{median}} or \textit{stepwise\_agg\textsubscript{mean}}, respectively; and (3) aggregation using a weighted mean approach, denoted as \textit{weighted\_agg}. This method weights each genus according to its number of species in the trait datasets, regardless of whether information was available for every used grouping feature (Figure \ref{fig:data_proc_overview}).

\subsubsection*{Comparison of family-level aggregated traits with family-level assigned traits}

To evaluate the effect of the five trait aggregation methods
(\textit{direct\_agg\textsubscript{median}}, \textit{direct\_agg\textsubscript{mean}}, 
\textit{stepwise\_agg\textsubscript{median}},
\textit{stepwise\_agg\textsubscript{mean}},
\textit{weighted\_agg}) on the assignment of trait affinity scores, we compared our aggregated affinities with trait affinities assigned at family level by experts. Expert-assigned trait data were available for a subset of grouping features and taxa for Australia (\cite{chessman_dissolved-oxygen_2018}) and North America (CONUS database), so we conducted this comparison only for these two datasets. For the Australian dataset, we used the grouping features feeding mode (all traits listed in Table 1 except parasite) and size (220 families $*$ 8 family–trait combinations = 1760 comparisons). For the North American dataset, we used the grouping features feeding mode, respiration, size, voltinism, and locomotion (94 families $*$ 17 family–trait combinations = 1598 comparisons). Assigned traits from North America were on a categorical scale, and we converted them to binary prior to comparing them with aggregated trait affinities. Size was assigned as continuous variable for Australia, thus, we converted size into three traits (small, medium, and large) with a binary coding. Because trait affinities ranged from 0 to 1, the maximum difference possible in trait affinity was 1 to –1, either which corresponds to 100 \%. For convenience and to improve interpretation, we calculated absolute trait differences.

\subsubsection*{Effect of taxonomic hierarchy and trait variability on aggregation outcomes}

To examine the influence of taxonomic hierarchy and trait variability on the outcomes of the different trait aggregation methods, we created three hypothetical scenarios. We simulated three different families, each composed of 25 total species but with different taxonomic hierarchical structures.  The three families consisted of (1) a family with an equal number of genera and species (five genera each with five species each), denoted as \textit{sim\_base}; (2) a family in which one genus had a much larger number of species than the other four genera (1 genus with 13 species, 4 genera with 3 species each), denoted as \textit{sim\_extreme}; (3) a family in which all genera had a different number of species (8, 2, 7, 3, 5), denoted as \textit{sim\_variation}. We assigned a hypothetical grouping feature with three traits (T1, T2, and T3) to each scenario. We then simulated the 25 affinities, one per species, for each trait by sampling from a truncated normal distribution bound by 0 and 1 and with a mean value of 0.5. To simulate different levels of trait variability, we repeated the sampling 100 times for each of the 5 levels of standard deviations (0.2, 0.4, 0.6, 0.8, and 1), resulting in 12,500 simulated trait affinities for each simulated trait ($25$ species per family $*$ 5 levels of trait variability $*$ 100 replicates). We converted simulated trait affinities to proportions, as was done during trait database processing, and assigned the 12,500 simulated affinities per trait to each of the three family scenarios. We then applied the five trait aggregation methods described above to each simulated dataset. We compared the resulting ranges of aggregated trait affinities between levels of trait variability and taxonomic hierarchical scenario as well as the differences in trait affinities obtained by each aggregation method.

\begin{figure}
  \centering
  \subfile{Flowchart/Flowchart_methods.tex}
  \caption{Data processing steps of the selected traits. Intermediate (gray) and main (orange) steps of data preparation are depicted. The dashed bottom box illustrates the different trait aggregation methods (hypothetical data in the upper left corner). The aggregation methods (purple) and intermediate steps of the aggregation methods (pink) are displayed. Abbreviations: EU, Europe; NOA, North America; AUS, Australia; NZ, New Zealand.}
  \label{fig:data_proc_overview}
\end{figure}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\newpage

\subsubsection*{Comparison of family-level aggregated traits with family-level assigned traits}

Aggregated trait affinities of the five trait aggregation methods (\textit{direct\_agg\textsubscript{median}}, \textit{direct\_agg\textsubscript{mean}}, \textit{stepwise\_agg\textsubscript{median}}, \textit{stepwise\_agg\textsubscript{mean}}, and \textit{weighted\_agg}) were compared to trait affinities assigned at family-level by experts, which were available for Australia and North America for a subset of grouping features and taxa. For the Australian dataset, we compared aggregated trait affinities with assigned trait affinities resolved at family-level for the grouping features feeding mode and size using data from \citet{chessman_dissolved-oxygen_2018}. In \citet{chessman_dissolved-oxygen_2018} feeding mode is classified similarly as in the Australian dataset except that the trait parasite is missing. We conducted the comparison for the 220 families available in \citet{chessman_dissolved-oxygen_2018}. Considering each factor combination of family and trait (in total 8) this amounted to 1760 cases.

For the North American dataset, we compared aggregated trait affinities with assigned trait affinities on family-level for the grouping features feeding mode, respiration, size, voltinism, and locomotion. The assigned trait affinities at family-level are part of the CONUS database and originate from expert knowledge. Trait information was available for 94 families of which all were present in the North American dataset (total number of cases 1598). The traits were on the categorical scale and were converted to binary traits prior to the comparison with aggregated trait affinities.

As mentioned above, trait affinities ranged from 0 to 1. Hence, the maximum difference possible in trait affinities is 1 or -1 (corresponds to 100 \%). For convenience and to improve interpretation, we report absolute trait differences.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\subsection*{Effects of harmonisation and trait aggregation on trait-environment relationship inferences}

To further evaluate the effects of grouping-feature harmonisation and trait aggregation on characterisation of trait–environment relationships, we repeated the analysis of \citet{szocs_effects_2014}, who studied the effects of anthropogenic salinisation on invertebrates in the River Werra in Germany. The River Werra has been subject to effluents from the potash industry since the mid- 20th century, which makes it a useful system for studying the responses of invertebrates and their trait compositions to salinisation (\cite{bathe_biological_2011}). \citet{szocs_effects_2014}  used redundancy analysis (RDA) to compare trait compositions among sites downstream, upstream, and close to salt discharge (transition). Further details about the original study can be found in \citet{szocs_effects_2014}.

We re-analysed the \citet{szocs_effects_2014} study by performing a RDA with harmonised grouping features from the dataset for Europe and additionally with traits that we aggregated to family level with the five aggregation methods described above. We used all 21 grouping features from the original study, but replaced 6 with harmonised grouping features from our European trait dataset, namely: size, feeding mode, locomotion, oviposition, respiration, and voltinism.
We compared the results of the original study’s RDA with the results of the RDA from the harmonised grouping features. Specifically, the trait composition, expressed as community weighted mean (CWM) traits, was ordinated along an electric conductivity gradient. We compared the species scores obtained from the RDA (i.e., the coordinates of the tips of the vectors representing the CWM traits in the bi- or triplots). Following the original study, we used the Mahalanobis distance measure and identified traits associated with higher or lower salinity based on their distance to the ordination axis median. We considered CWM traits with a Mahalanobis distance greater than the 97.5 \%-quantile of the Chi-square distribution (5.02) to have responded to either lower or higher salinity. Additionally, for testing the effect of different aggregation methods on inference about CWM trait–salinisation relationships, we repeated the RDA five times, each with family-level aggregated trait values assigned to corresponding taxa from \citet{szocs_effects_2014}.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\subsection*{Data analysis}

All data processing and RDA analyses were carried out using R (\cite{cite_R}). RDA was computed using the vegan package (\cite{cite_vegan}). Raw data and the R code for data processing and grouping feature harmonisation are located in the Github repository: \url{https://github.com/KunzstLD/Invertebrate_traits}. Scripts and data to reproduce the trait aggregation, comparison of aggregated and assigned traits, and the RDA analysis are located in the Github repository \url{https://github.com/KunzstLD/Trait-aggregation}.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section*{Results}

\subsection*{Harmonised trait datasets}

\subsection*{Taxonomic coverage of the harmonised trait datasets}

The harmonised European, North American, Australian, and New Zealand trait datasets differed in terms of their taxonomic coverage. The European trait dataset had the largest taxon pool with 4601 taxa (Table \ref{tab:tax_coverage}) followed by the North American dataset (3753 taxa), the Australian dataset (1402 taxa), and the New Zealand dataset with the smallest taxon pool (478 taxa). The European, New Zealand, and North American datasets contained most taxa at the species level, whereas the Australian dataset comprised a similar number of taxa at species and genus level.

\begin{table}[ht]
    \centering
    \caption{Number (Nr.) of taxa per harmonised dataset and per taxonomic level. Numbers in parenthesis show rounded relative frequencies in percent.} 
    \label{tab:tax_coverage}
    \begin{tabular}{lccccc}
    \toprule[.1em]
    Dataset & Taxa (Nr.) & Aquatic insects (Nr.) & Species & Genus & Family \\ 
    \toprule[.1em]
    EU & 4601 & 3942 (86) & 3739 (81) & 704 (15) & 158 (3) \\ 
    NOA & 3753 & 3305 (88) & 2414 (64) & 1163 (31) & 176 (5)  \\ 
    AUS & 1402 & 1016 (72) & 564 (40) & 578 (41) & 260 (19) \\ 
    NZ & 478 & 443 (93) & 404 (85) & 47 (10) & 27 (6) \\ 
    \bottomrule
    \end{tabular}
\end{table}
\begin{minipage}{\linewidth}{\fontsize{8}{10}\selectfont
\centering
Abbreviations: EU, European; NOA, North American; AUS, Australian; NZ, New Zealand.
\end{minipage}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\subsection*{Completeness of trait information}

The numbers of entries with available information for the selected grouping features varied considerably within the harmonised European, North American, and Australian datasets (Table \ref{tab:trait_coverage}). By contrast, the New Zealand dataset contained complete trait information for most of the investigated grouping features (94–100\%).

The amount of entries with available information for the selected grouping features varied strongly for the developed European, North American, and Australian datasets (Table \ref{tab:trait_coverage}). By contrast, the New Zealand dataset contained complete trait information for most of the investigated grouping features (between 94 \% and 100 \%).

\begin{table}[ht]
    \centering
    \caption{Rounded percentage of entries that include 
    information for the individual grouping features
    shown per trait dataset.} 
    \label{tab:trait_coverage}
    \begin{tabular}{lccccccc}
    \toprule[.1em]
    Dataset & Body form & Oviposition & Voltinism & Locomotion & Size & Respiration & Feeding mode \\ 
    \toprule[.1em]
    EU & 8 & 15 & 23 & 36 & 11 & 57 & 76 \\ 
    NOA & 28 & 13 & 47 & 52 & 73 & 44 & 63 \\ 
    AUS & 4 & 46 & 49 & 39 & 75 & 68 & 99 \\ 
    NZ & 100 & 94 & 100 & 99 & 100 & 100 & 99 \\ 
    \bottomrule
    \end{tabular}
\end{table}
\begin{minipage}{\linewidth}{\fontsize{8}{10}\selectfont
\centering
Abbreviations: EU, European; NOA, North American; AUS, Australian; NZ, New Zealand.
\end{minipage}

\newpage

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\subsection*{Discrepancies in invertebrate trait definitions across databases}

Definitions of grouping features and traits varied in their level of detail in the investigated trait databases. The freshwaterecology.info, Tachet, and CONUS databases provided more detailed descriptions of trait information than the Vieira and New Zealand databases provided. The Australian trait database (\cite{kefford_integrated_2020}) was a collection of seven specific trait datasets; thus, grouping features occurred multiple times with varying differentiation into traits. Depending on the dataset, trait information was described with more or less detail.

The definitions of grouping features also varied across databases in their differentiation into traits and in their focus, both which can lead to discrepancies in trait definitions. We provide a summary of discrepancies in trait definitions in the appendix (Table \ref{stab:trait_definitions}). Varying levels of differentiation among the trait databases were present in all investigated grouping features (Table \ref{tab:trait_databases_coding_differentiation} and Table \ref{stab:trait_definitions}).

For example, for the grouping feature feeding mode, discrepancies arose because traits were assigned in different ways. The Tachet database defined predators as carvers, engulfers, and swallowers. By contrast, the CONUS database defined predators as engulfers and carnivorous piercers. In turn, the Tachet database defined piercers as a separate trait encompassing both herbivorous and carnivorous piercers. Furthermore, the focus in the freshwaterecology.info database for feeding mode was primarily on the food source of a species (except for filterers), whereas the other databases focused on the strategies of food acquisition. For example, the freshwaterecology.info database defined predator as "eating from prey", and the other databases used the mouthpart morphology as basis for their definitions. The Tachet database captured the food source in an additional grouping feature. Locomotion definitions also differed in focus among databases. The New Zealand and freshwaterecology.info databases described locomotion as an organism’s way of movement, Tachet as substrate relation and locomotion, Vieira as how organisms deal with flow, Australia as attachment, and CONUS included not only the method of movement but also the location of movement. Similarly, databases differed in their focus for reproduction traits. In the freshwaterecology.info and Tachet databases, reproduction was captured in one grouping feature and was defined as location of oviposit clutches and mode of reproduction. The Vieira database provided information on the oviposition location but not on reproductive behaviour. The Australian database reported grouping features for reproductive behaviour but also on oviposition location separately. The New Zealand database distinguished three grouping features related to reproduction: reproductive technique, oviposition location (e.g. water surface, terrestrial), and egg/egg mass (e.g. free, cemented).

Trait affinity scores also varied among the databases and were described by varying coding methods (e.g., binary, fuzzy, continuous). In the freshwaterecology.info and Australian databases a combination of coding methods is used, whereas in the Tachet and New Zealand databases exclusively fuzzy coding is used. The Vieira and CONUS databases contained categorical grouping features that can be converted into binary trait affinities (Table \ref{tab:trait_databases_coding_differentiation}).

\begin{landscape}
\begin{longtable}{m{2.3cm}|m{3cm}|m{1.8cm}|m{1.4cm}|m{1.4cm}|m{4.1cm}|m{2cm}}
\caption{Number of traits per grouping feature and type of coding of the traits for the grouping features used in this study per database. Oviposition location was used for the New Zealand database.}
\endfirsthead
\toprule[.1em]
\label{tab:trait_databases_coding_differentiation}
\specialcell{Grouping \\ feature} & \specialcell{freshwater- \\ ecology.info} & Tachet & CONUS & Vieira & Australia & New Zealand \\
\toprule[.1em]
Feeding Mode & \begin{tabular}[c]{@{}l@{}}10 traits; \\ 10 point assginment \\ system\end{tabular}       & \begin{tabular}[c]{@{}l@{}}7 traits; \\ fuzzy {[}0 - 3{]}\end{tabular} & \begin{tabular}[c]{@{}l@{}}6 traits; \\ binary\end{tabular}   & \begin{tabular}[c]{@{}l@{}}8 traits; \\ binary\end{tabular}  & \begin{tabular}[c]{@{}l@{}}16 traits$^{\dagger}$; \\ binary, proportional {[}0 - 1{]}, \\ fuzzy {[}0 - 3{]}\end{tabular} & \begin{tabular}[c]{@{}l@{}}6 traits; \\ fuzzy {[}0 - 3{]}\end{tabular} \\
\midrule
Voltinism                                                           & \begin{tabular}[c]{@{}l@{}}6 traits; \\ single category \\ assignment system\end{tabular} & \begin{tabular}[c]{@{}l@{}}3 traits; \\ fuzzy {[}0 - 3{]}\end{tabular} & \begin{tabular}[c]{@{}l@{}}3 traits;\\ binary\end{tabular}    & \begin{tabular}[c]{@{}l@{}}3 traits; \\ binary\end{tabular}  & \begin{tabular}[c]{@{}l@{}}7 traits; \\ binary, proportional {[}0 - 1{]}, \\ fuzzy {[}0 - 3{]}\end{tabular}  & \begin{tabular}[c]{@{}l@{}}3 traits; \\ fuzzy {[}0 - 3{]}\end{tabular} \\
\midrule
Locomotion                                                          & \begin{tabular}[c]{@{}l@{}}6 traits; \\ 10 point assignment \\ system\end{tabular}        & \begin{tabular}[c]{@{}l@{}}8 traits; \\ fuzzy {[}0 - 5{]}\end{tabular} & \begin{tabular}[c]{@{}l@{}}10 traits; \\ binary\end{tabular}  & \begin{tabular}[c]{@{}l@{}}9 traits; \\ binary\end{tabular}  & \begin{tabular}[c]{@{}l@{}}9 traits; \\ binary, fuzzy  {[}0 - 3{]}\end{tabular}                              & \begin{tabular}[c]{@{}l@{}}4 traits; \\ fuzzy {[}0 - 3{]}\end{tabular} \\
\midrule
Respiration                                                         & \begin{tabular}[c]{@{}l@{}}7 traits; \\ binary\end{tabular}                               & \begin{tabular}[c]{@{}l@{}}5 traits; \\ fuzzy {[}0 - 3{]}\end{tabular} & \begin{tabular}[c]{@{}l@{}}3 traits;  \\ binary\end{tabular}  & \begin{tabular}[c]{@{}l@{}}8 traits; \\ binary\end{tabular}  & \begin{tabular}[c]{@{}l@{}}10 traits; \\ binary, proportional {[}0 - 1{]}, \\ fuzzy {[}0 - 3{]}\end{tabular} & \begin{tabular}[c]{@{}l@{}}4 traits; \\ fuzzy {[}0 - 3{]}\end{tabular} \\
\midrule
\begin{tabular}[c]{@{}l@{}}Reproduction/\\ Oviposition\end{tabular} & \begin{tabular}[c]{@{}l@{}}9 traits; \\ binary\end{tabular}                               & \begin{tabular}[c]{@{}l@{}}8 traits; \\ fuzzy {[}0 - 3{]}\end{tabular} & \begin{tabular}[c]{@{}l@{}}10 traits;  \\ binary\end{tabular} & \begin{tabular}[c]{@{}l@{}}10 traits; \\ binary\end{tabular} & \begin{tabular}[c]{@{}l@{}}13 traits$^{\ddagger}$; \\ binary\end{tabular}                                                 & \begin{tabular}[c]{@{}l@{}}4 traits; \\ fuzzy {[}0 - 3{]}\end{tabular} \\
\midrule
Size                                                                & -                                                                                         & \begin{tabular}[c]{@{}l@{}}7 traits;\\ fuzzy {[}0 - 3{]}\end{tabular}  & \begin{tabular}[c]{@{}l@{}}3 traits; \\ binary\end{tabular}   & \begin{tabular}[c]{@{}l@{}}3 traits; \\ binary\end{tabular}  & \begin{tabular}[c]{@{}l@{}}9 traits; \\ binary, continuous, \\ \\ fuzzy {[}0 - 3{]}\end{tabular}             & \begin{tabular}[c]{@{}l@{}}5 traits; \\ fuzzy {[}0 - 3{]}\end{tabular} \\
\midrule
Body Form                                                           & -                                                                                         & -                                                                      & -                                                             & \begin{tabular}[c]{@{}l@{}}4 traits; \\ binary\end{tabular}  & \begin{tabular}[c]{@{}l@{}}4 traits; \\ fuzzy {[}0 - 3{]}\end{tabular}                                       & \begin{tabular}[c]{@{}l@{}}4 traits; \\ fuzzy {[}0 - 3{]}\end{tabular} \\
\bottomrule
\end{longtable}
\begin{minipage}{\linewidth}\small
      $\dagger$ Some of the feeding mode traits used in the Australian database were similar (e.g. trait \textit{Shredder}, \textit{Shredder, Detrivore}, and \textit{Collector, Shredder}).
      \newline
      $\ddagger$ Not all traits were considered because trait information was partly presented as comments to describe other traits or due to incomplete information.
  \end{minipage}
\end{landscape}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\subsection*{Comparing aggregation methods}

\subsection*{Comparison of family-level aggregated traits with family-level assigned traits}
\label{sec:diff_trait_agg_chessman}

The comparison of family-level aggregated trait affinities with those assigned by experts for the Australian and North American databases showed that aggregation method (median or mean) had a greater effect than overall approach (direct, stepwise, or weighted) in whether trait affinities differed from those assigned by experts.

The percentage of differing cases of aggregated versus assigned trait affinities varied by method between 16.2 and 22.9 \% for the Australian dataset and between 15.3 and 47 \% for the North American dataset (Table \ref{tab:summary_stat_aggr_vs_fam_assigned}). In general, median aggregation methods yielded fewer cases with differences compared with mean and weighted aggregation methods. However, median aggregation methods produced, on average, greater absolute differences for both datasets. Standard deviations of absolute differences were similar for all tested aggregation methods. Maximum differences of 1 (i.e., the maximum possible difference) occurred for all investigated grouping features in both datasets (Figure \ref{fig:diff_aggr_traits_chessman} and Figure \ref{fig:diff_aggr_traits_pyne}).

\begin{table}[H]
  \centering
  \caption{Percentage of differing cases, minimum, maximum, mean, and standard deviation of absolute differences between trait affinities assigned at family level by experts and aggregated trait affinities from five different aggregation methods.}
  \label{tab:summary_stat_aggr_vs_fam_assigned}
  \begin{tabular}{ll|ccccc}
  \toprule[.1em]
  Database & \specialcell{Comparison to\\ traits at family level} & \specialcell{Differing \\ cases [\%]} & \specialcell{Min. \\ differences} & \specialcell{Max. \\ differences} & \specialcell{Mean abs. \\ differences} & \specialcell{SD abs. \\ differences} \\ 
  \toprule[.1em]
  \multirow{4}{*}{\specialcell{Australia}} & \specialcell{\textit{direct\_agg (median)}} & 16.53 & 0.01 & 1.00 & 0.45 & 0.27 \\ 
  & \specialcell{\textit{direct\_agg (mean)}} & 23.24 & $< 0.01$ & 0.99 & 0.34 & 0.23 \\ 
  & \specialcell{\textit{stepwise\_agg (median)}} & 17.90 & 0.01 & 1.00 & 0.42 & 0.26 \\ 
  & \specialcell{\textit{stepwise\_agg (mean)}} & 23.24 & $< 0.01$ & 0.99 & 0.33 & 0.22 \\ 
  & \specialcell{\textit{weighted\_agg}} & 23.24 & $< 0.01$ & 1.00 & 0.34 & 0.24 \\ 
  \midrule
  \multirow{4}{*}{\specialcell{North America}} & \specialcell{\textit{direct\_agg (median)}} & 15.33 & 0.17 & 1.00 & 0.70 & 0.26 \\ 
  & \specialcell{\textit{direct\_agg (mean)}} & 47.00 & $< 0.01$ & 1.00 & 0.30 & 0.26 \\ 
  & \specialcell{\textit{stepwise\_agg (median)}} & 18.00 & 0.08 & 1.00 & 0.63 & 0.28 \\ 
  & \specialcell{\textit{stepwise\_agg (mean)}} & 47.00 & $< 0.01$ & 1.00 & 0.30 & 0.27 \\ 
  & \specialcell{\textit{weighted\_agg}} & 47.00 & $< 0.01$ & 1.00 & 0.31 & 0.28 \\ 
  \bottomrule
  \end{tabular}
\end{table}

\newpage

\begin{figure}[H]
  \centering
  \includegraphics[width=16.5cm, height=10cm]{Deviances_trait_agg_chessman.png}
  \caption{Cases (factor combination of investigated families and traits) where differences occurred between aggregated traits and expert assigned traits at family level for the Australian dataset for the grouping features feeding mode and body size. Violin plots - mirrored density plots - show
  the density of the absolute trait affinity differences. Absolute differences in trait affinities are depicted as grey dots. \textit{n} denotes the number of cases per comparison where differences occurred. The blue dot indicates the mean of absolute differences, and the error bars show the standard deviation. The grey vertical lines show the 25\textsuperscript{th}, 50\textsuperscript{th} and 75\textsuperscript{th} quantiles of the density estimate.}
  \label{fig:diff_aggr_traits_chessman}
\end{figure}

\begin{figure}[H]
  \centering
  \includegraphics[width=16.5cm, height=10cm]{Deviances_trait_agg_pyne.png}
  \caption{Cases (factor combination investigated families and traits) where differences occurred between aggregated traits and traits assigned at family-level. Violin plots - a mirrored density plot - show the density of the absolute trait affinity differences for the North America dataset for the grouping features feeding mode, locomotion, respiration, body size and voltinism. For more details see Figure \ref{fig:diff_aggr_traits_chessman}.}
  \label{fig:diff_aggr_traits_pyne}
\end{figure}

\newpage

\subsection*{Simulation of varying taxonomic hierarchies}

We evaluated the simulation results based on the produced range of aggregated trait affinities and by comparing results of every simulated dataset.

As expected the produced range of trait affinities increased for all aggregation methods with increasing trait variability (Figure \ref{fig:overview_sim_results}).% Ben would add: 
 % as indicated by the increasing range of the boxes and wiskers with increasing standard deviation.

 % Ben further suggests: When each genus has the same number of species (sim base) ...... . In contrast with most species in one genus and fewer in the other genera (sim_extreme) .... . While with variable number of species per genera () ..... .

For the \textit{sim\_base} scenario, the mean aggregation methods yielded similar ranges of aggregated trait affinities within each level of trait variability whereas the median aggregation methods produced mostly greater ranges of aggregated trait affinities. By contrast, the ranges of trait affinities slightly differed for all aggregation methods for the simulation scenarios \textit{sim\_extreme} and \textit{sim\_variation}. \textit{Weighted\_agg} and \textit{stepwise\_agg\textsubscript{mean}} produced a wider range of trait affinities than the \textit{direct\_agg\textsubscript{mean}} in the \textit{sim\_extreme} scenario. For most levels of trait variability the \textit{stepwise\_agg \textsubscript{median}} resulted in the largest range of trait affinities in the \textit{sim\_extreme} and \textit{sim\_variation} scenarios.

Each trait aggregation method yielded one aggregated affinity for every simulated dataset, in total 15000 comparisons ($100$ replicates $*$ 5 levels of trait variability $* 10$ unique comparisons of trait aggregation methods $* 3$ scenarios). In most simulated datasets the different aggregation methods often resulted in similar trait affinities. Only 1.42 \% (213 out of 15.000) of all comparisons showed a difference equal or greater than an absolute trait affinity of 0.1. Most of these differences occurred for the \textit{sim\_extreme} scenario (83.5 \%). The majority of the differences equal or above 0.1 were found between the aggregation methods \textit{direct\_agg\textsubscript{mean}} and \textit{stepwise\_agg\textsubscript{median}}, \textit{direct\_agg\textsubscript{median}} and \textit{stepwise\_agg\textsubscript{median}}, and \textit{stepwise\_agg\textsubscript{median}} and \textit{weighted\_agg} (Figure \ref{fig:sim_indv_runs}). 
% Differences were rather small, largest difference 0.265 for sim_extreme (stepwise_median vs weighted_agg)

\begin{figure}[H]
  \centering
  \includegraphics[width=16.5cm, height=10cm]{Overview_sim_results.png}
  \caption{Ranges of aggregated trait affinities for the three examples of taxonomic hierarchies and simulated levels of trait variability. Boxplots depict results for each trait aggregation method of 100 simulations. T1, T2, and T3 are the simulated traits.}
  \label{fig:overview_sim_results}
\end{figure}

\begin{figure}[H]
  \centering
  \includegraphics[width=16.5cm, height=10cm]{Diffs_indiv_runs_sim.png}
  \caption{Comparison of the aggregated trait affinities produced by the different trait aggregation methods for every simulated dataset across all 3 simulated traits. Dots depict comparisons where absolute differences between aggregated trait affinities were greater than 0.1. \newline
  Comparison: \newline
  1) Direct\_agg (median) - Stepwise\_agg (mean) \newline
  2) Direct\_agg (median) - Weighted\_agg, \newline
  3) Stepwise\_agg (mean) - Stepwise\_agg (median), \newline
  4) Stepwise\_agg (mean) - Weighted\_agg, \newline  
  5) Direct\_agg (mean) - Stepwise\_agg (median), \newline
  6) Direct\_agg (median) - Stepwise\_agg (median), \newline
  7) Stepwise\_agg (median) - Weighted\_agg \newline
  }
  \label{fig:sim_indv_runs}
\end{figure}

\newpage

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\subsection*{Effects of harmonisation and trait aggregation on trait-environment relationships}

The original RDA in \citet{szocs_effects_2014} indicated that downstream sites (higher salinity) were characterised by the traits shredder, ovoviviparity, multivoltinism, long life cycle duration ($> 1$ year), and gill respiration, whereas upstream sites (lower salinity) were characterized by univoltinism, oviposition in clutches and short life cycle duration ($< 1$ year).

Overall, harmonised trait data and all aggregation methods produced similar distributions of RDA species scores (Figures \ref{fig:violin_plot_species_sc}) compared to the original analysis in \citet{szocs_effects_2014}. Using harmonised grouping features resulted in less traits that distinguish upstream and downstream sites in comparison to the original analysis (Figures \ref{fig:violin_plot_species_sc} and \ref{fig:boxplots_scores_on_constrained_axis}). According to the RDA of the trait composition, downstream sites were characterised by taxa with the traits multivoltinism, ovoviviparity and long life cycle duration. Upstream sites were characterised by univoltine taxa with a short life cycle duration that lay their eggs in an aquatic environment (aquatic eggs). The trait aquatic eggs of the harmonised grouping feature oviposition was derived by amalgamating the trait oviposition in clutches and other related traits (Table \ref{tab:traits_harmonisation}). The traits shredder and gill respiration did not characterise sites with higher salinisation.

Using traits aggregated at family-level from the harmonised dataset also showed results similar to the original analysis, but with less traits distinguishing upstream and downstream sites (Figures \ref{fig:violin_plot_species_sc} and \ref{fig:boxplots_scores_on_constrained_axis}). The \textit{direct\_agg\textsubscript{mean}}, \textit{direct\_agg\textsubscript{median}}, and \textit{weighted\_agg} 
characterised the downstream sites with the same traits as the original analysis except that downstream sites were not characterised by the trait shredder. Upstream sites were characterised by the traits univoltinism, aquatic eggs and short life cycle duration. The same results were obtained when re-analysing the data with traits aggregated by the \textit{stepwise\_agg\textsubscript{mean}} and \textit{stepwise\_agg\textsubscript{median}}, with the exception that none of the life cycle traits characterised upstream or downstream sites. Thus, in our comparison aggregating by \textit{direct\_agg\textsubscript{mean}}, \textit{direct\_agg\textsubscript{median}}, and \textit{weighted\_agg} yielded the least change in interpretation of the RDA results compared to the original findings.

\begin{figure}[H]
    \centering
    \includegraphics[width=16.5cm, height=10cm]{Species_scores_rda.png}
    \caption{Species scores obtained by RDA from the original analysis (\cite{szocs_effects_2014}), using harmonised grouping features, and using harmonised grouping features with traits aggregated to family-level. The dots represent the individual species scores for each analysed trait along the conductivity axis. The violin plot shows the density estimate of the species scores. Gray vertical lines indicate the median of the obtained species scores.}
    \label{fig:violin_plot_species_sc}
\end{figure}

% Old figure:
% \begin{sidewaysfigure}
%   \label{fig:boxplots_scores_on_constrained_axis}
%   \centering
%   \includegraphics[width= 22cm, height=12cm]{Scores_on_constrained_axis_combined.pdf}
%   \caption{RDA of traits constrained by electric conductivity for the tested methods and the original study. Shown are boxplots of the site scores along the conductivity axis. The rug on the right side of each plot indicates species scores of the traits on the conductivity axis. Only traits with a mahalanobis distance greater than the $97.5 \%$ quantile of the Chi-square distribution (5.02) were labelled.}
% \end{sidewaysfigure}

\begin{figure}[H]
  \centering
  \includegraphics[width=16.5cm, height=25cm]{Scores_on_constrained_axis_comb_ALT.pdf}
\end{figure}
\clearpage % insert a page break
\captionof{figure}{RDA of traits constrained by electric conductivity for the tested methods and the original study. Shown are boxplots of the site scores along the conductivity axis. The rug on the right side of each plot indicates species scores of the traits on the conductivity axis. Only traits with a Mahalanobis distance greater than the $97.5 \%$ quantile of the Chi-square distribution (5.02) were labelled. For better comparability, species scores of the original analysis were multiplied by -1.} 
\label{fig:boxplots_scores_on_constrained_axis}

\newpage

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section*{Discussion}

% From Freshwater biology: 
% This should highlight the significance of the results and place them in the context of other work. It should not include references to figures or tables cited in the Results – check that all of your Results are presented in the Results section. Figures (e.g. photographs, conceptual diagrams) or tables presented specifically as part of the Discussion are welcome and may be cited there.

\subsection*{Trait definition discrepancies and taxonomic resolution}

Synthesizing trait information from multiple trait databases is crucial for developing the full potential of trait-based approaches, e.g. for studying community trait responses to environmental gradients worldwide. Our attempt at harmonising invertebrate trait information from trait databases of different regions showed that harmonisation was labour-intensive because grouping features are resolved differently into traits and various codings are used to describe traits in the various databases. In addition, the same traits are sometimes defined differently, requiring expert knowledge to reclassify these traits (e.g. the trait piercer in the Tachet database). Although, others have noted the lack of standardised trait terminology in freshwater ecology (\cite{baird_toward_2011, brink_traits-based_2011}), we are the first to describe invertebrate trait definition discrepancies for some commonly used grouping features. To resolve definition discrepancies and facilitate data synthesis in the future, terminological standards are needed. Harmonised definitions and concepts of traits have been developed in the past for other organism groups, e.g. for plants with the \textit{Thesaurus Of Plant characteristics} (TOP) initiative (\cite{garnier_towards_2017}). The core of this harmonisation initiative is to provide standardised trait definitions and to draw connections to synonyms, related terms, and surrounding concepts by linking to other controlled vocabularies or ontologies in the field. Such initiatives could be a role model for freshwater ecologists to establish unambiguous terminologies for invertebrate traits. The existing freshwater invertebrate trait databases could be linked through standardised terminologies or ontologies, as suggested by \citet{baird_toward_2011} a decade ago. By following the recently proposed \textit{Ecological Trait-data Standard Vocabulary} (ETS), providers of invertebrate trait data could connect their traits to such ontologies (\cite{schneider_towards_2019}). Once standardised trait definitions are established, these will improve trait data sharing, trait data processing when working with multiple trait databases, and ultimately interpretation of the derived results.

Our analysis aimed to use most of the available invertebrate trait information for different regions to establish harmonised grouping feature datasets. Although many taxa are covered in the trait databases from which we extracted our datasets, the availability of trait information varies strongly across grouping features and taxonomic resolution differs between databases. Much information is available for grouping features that are often used in trait-environment studies, such as feeding mode and respiration. Surprisingly, there is little information on body form, a trait that is relatively easy to determine.

While trait information from Europe and New Zealand was mainly at species-level, considerable portions of trait data from North America and Australia were at genus or family-level. Assigned traits at family-level may not reflect the real trait diversity within a taxon, for example for the large, ecologically diverse family Chironomidae (\cite{serra_synthesising_2016}). However, species-level identification can be complicated, time-consuming and expensive (\cite{marshall_taxonomic_2006, resh_which_2008}). Consequently, trait information at family-level has been widely used by freshwater ecologists, for example in bioassessment (\cite{beketov_spear_2009}).

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\subsection*{Trait aggregation}

When trait values are only available at species-level and observational data are at less precise taxonomic levels, various trait aggregation methods have been used by researchers. We aggregated traits from an Australian and a North American trait dataset to family-level using 5 different aggregation methods and compared the results to traits originally assigned at family-level for these regions. Evaluation of the differences between aggregated and assigned traits is difficult because it remains unclear what the true value of a particular trait for a particular family is. Aggregation of trait information at species or genus-level to a point estimate at family-level suggests a precision that is not necessarily present, especially not for traits with a high variability or if trait information at species-level is missing. Some traits can vary strongly at more precise taxonomic levels than the family-level. For example, \citet{monaghan_improving_2013} found a high intra-family trait dispersion in the Tachet database for traits of the grouping features body size, flow preference, and reproduction. Further, studies focusing on the lability of traits, i.e. how much traits are constrained by phylogeny, found that traits of ecological preferences (e.g. thermal preference), body size, resistance forms, and to a lesser extent feeding mode are labile, and thus possibly highly variable (\cite{poff_functional_2006, wilkes_traitbased_2020}). If trait aggregation is necessary, however, our study can give guidance, under which circumstances the choice of aggregation method may be important. The 5 aggregation methods we tested use either the mean or the median and different weightings. Our results indicated that 1) the median aggregation methods compared to mean aggregation methods were often closer to the assigned traits at family-level and 2) the different weightings tested exerted a minor impact on the aggregated trait affinities. We obtained for both datasets more strongly differing cases for mean aggregation methods, irrespective of the weighting. In the Australian dataset, the amount of differing cases was slightly higher for the mean aggregation methods (23 \%) than for median aggregation methods (16 to 18 \% differing cases). In the North American dataset, mean aggregation methods yielded 47 \% cases being different to the assigned traits, a value much higher in comparison to the median aggregation methods (between 15 and 18 \%). When differences occurred between median aggregation methods and assigned traits, these differences were greater compared to the mean aggregation methods, particularly for the North American dataset. There, the mean absolute differences for the median aggregations were twice as high (0.63 - 0.7) compared to the mean aggregations (0.3 - 0.31). These differences can be explained by the binary coding used in the North American trait dataset and by the assigned traits. As a result, traits have affinities of either 1 or 0. Therefore, traits of a particular grouping feature showed in most cases a higher difference in trait affinities to each other than if they were fuzzy coded. Using the median to aggregate binary coded taxa often resulted in a value of 0 or 1, while using the mean yielded values between 0 and 1. Thus, using the median led for the North American dataset either to agreement with the assigned traits or to higher differences compared to the mean aggregation methods.

The weightings explored in this analysis should lead to different aggregated trait affinities in cases where genera within a family have different numbers of species. In fact, the simulation results indicated that for a particular family where one genus has a much larger number of species compared to the other genera, weighting the number of taxa can lead to different aggregated trait affinities, especially for traits with higher levels of variance. However, in the comparison of the trait aggregation methods to assigned traits the number of differing cases and the mean absolute differences in trait affinities were similar across the mean aggregation and across the median aggregation methods, which suggests a small influence of the weightings on the aggregation methods. Also, the distributions of absolute trait affinity differences to assigned traits were relatively similar for the mean aggregation methods and the median aggregation methods, respectively. The minor impact on trait aggregation of the weightings may be explained by the fact that a considerable portion of taxa had low numbers of genera or species. Of the taxa that were compared from the North American trait dataset, 14 \% were identified at family and 62 \% at genus-level, 52 \% comprised 5 or fewer genera, and 13 \% contained just one genus (Figure \ref{fig:tax_hierarchy_NOA}). In the Australian dataset, 21 \% of the compared taxa were identified at family, 40 \% at genus, and 39 \% at species-level, 68 \% of the taxa contained 5 or fewer genera, and 40 \% just one genus (Figure \ref{fig:tax_hierarchy_AUS}). Hence, these results could change when more trait information at species-level becomes available.

\subsection*{Grouping feature harmonisation}

We explored how using harmonised grouping features and aggregated traits might change the results of trait-environment relationships. We are not aware of other studies that compared trait-environment relationships using harmonised and non harmonised data. The re-analysis of the data on salinisation effects on biological traits from \citet{szocs_effects_2014} yielded only slightly different results compared to the original study, i.e. some of the traits that had responded in the original analysis did not respond in the re-analysis. These non-responsive traits were those closest to the criterion that defined when a trait was associated with either higher or lower salinity, namely feeding mode shredder, respiration gills, and life cycle duration traits. Feeding mode shredder was the only trait that was non-responsive when using harmonised (but not aggregated) as well as when using aggregated data. This result was likely due to the harmonisation procedure, when this trait was amalgamated based on three traits (miner, xylophagus, and shredder). Consequently, the trait affinities in the original data had a higher mean and standard deviation than in the aggregated and harmonised data (Table \ref{tab:SI_resp_traits_summary_stats}), suggesting that the signal in the original data was weakened by the harmonisation. %This might be the reason that shredders showed a less strong reaction to the salinity gradient in the re-analysis. -> Kicked out by CM
These findings show that if harmonisation is necessary, harmonised and non-harmonised data, if available, should be compared and possible averaging effects should be considered in further analyses. The fact that the family-level aggregated traits used in the re-analysis showed similar results compared to the original analysis agrees with previous findings that family-level traits can be sensitive enough to detect environmental impacts (\cite{beketov_spear_2009}). 
% Cite another paper that uses family-level trait data
% Although species-level traits are expected to result in more precise results when investigating trait-environment relationships.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\subsection*{Outlook}

Although comprehensive freshwater invertebrate trait databases have been developed for several regions in the past, data synthesis is difficult because of discrepancies in trait definitions. By providing an overview of these definition discrepancies we have set a starting point for the development of standardised trait terminology through which invertebrate trait databases can be linked. A consent on standard terminology and the subsequent development of ontologies are the next steps to facilitate trait-based analysis at large geographical scales. As our analysis showed, some grouping features might need to be re-classified to fit into such a standardised terminology. We could show that trait affinities resulting from fuzzy coding and binary coding can be used together, but a uniform coding of traits is another problem that should be addressed during trait standardisation.

As ecological knowledge increases in the future, and with the aid of new  methods such as computer vision for species and trait identification (\cite{hoye_deep_2020}), more observational and trait data at species-level may become available in the future, reducing the need for trait aggregation. However, if traits need to be aggregated we showed that trait aggregation is often in agreement with expert assignments, especially aggregation approaches using the median. Ultimately, the question which aggregation method is the objectively best measure cannot be answered. Nevertheless, we make the following recommendations for situations when trait aggregation is necessary. 1) The aggregation method will tend to have little effect except for highly variable traits and taxa where a few genera have a high number of species. 2) The coding used to describe traits (binary, fuzzy) should be considered, because median and mean aggregations can yield very different results for binary coded traits. 3) Weighting trait information based on the taxa present in the database used does not seem to have much influence except for families with few genera that have a high number of species. 4) A measure of trait dispersion should be reported to indicate the uncertainty of the aggregated estimate. 
% OLD: No matter which aggregation approach is chosen, for future studies in which researchers must aggregate traits,....

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Further ideas for the discussion:
% - What do we really know because someone has watched the animals and published their behaviour; what is extrapolated knowledge? 
% - Where does the classification come from? Few sources used in NOA also used in AUS:
% AUS: Merrit \& Cummins 1978/1984 for feeding mode for a few genera and families 
% NOA: Merritt, Richard W., Cummins, Berg 2008 -> but not clear if for feeding mode as well; according to trait description "Cummins, K. W. Trophic Relations of Aquatic Insects" was used for feeding mode
% - Discussion on fuzzy coding -> Uncertainty captured by fuzzy codes? 
% - Extrapolate to "large" scale studies, i.e. studies with a stronger gradient?
% - Traits that responded differently after harmonisation:
% Resp gills: Original data has a lot of zeros, which are not present in the harmonised \& aggregated datasets. In the analysis with the harmonised dataset resp gills is not responding to the salinity gradient. Aggregated data look similar to the harmonised ones, but resp gills is here one of the traits responding to conductivity!
% Why did some aggregated datasets promote life cycle duration traits and others not? Results using trait data obtained when aggregating with stepwise\_agg (median) and stepwise\_agg (mean) indicate that life cycle duration traits are not characterising upstream or downstream sites.
% Results using trait data obtained with the other aggregation methods indicate that life cycle duration traits characterise upstream or downstream sites
% However: Mean, median and SD of these traits are all relatively similar across all aggregated datasets (see Table S2). 
% Include Ben's point? Discussion on evolutionary history?
% Discussion point BK: It is intersting that Australia has fewer species and genus than NOA and EU (not suprisngly since the datasets in Oz were mostly delveoped at family level) but more families than these other regions. I wonder if this indicates that Family richness is really higher in Australia. Perhaps it might be as Australia has a wider range of evolutionary histories. This would be worth considering in the discussion

\newpage

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% Bibliography %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\printbibliography

\newpage

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% Supporting information %%%%%%%%%%%%%%%%%%%%%%%%
\setcounter{table}{0}
\setcounter{figure}{0}
\renewcommand{\thetable}{S\arabic{table}}
\renewcommand{\thefigure}{S\arabic{figure}}

\subfile{Results/SI}

\end{document}